---
title: "Appendix 2"
output: pdf_document
date: "2025-11-18"
toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, messages = FALSE, warning = FALSE)

library(tidyverse)
library(sf)
library(nimble)
# remotes::install_github("rileymummah/SpFut.flexiSDM", build_vignettes = T)
library(SpFut.flexiSDM)
```

<!-- ## --------------------------- -->
<!-- ## Objective: -->
<!-- ##    - Imports species and covariate data -->
<!-- ##    - Scales covariates -->
<!-- ##    - Sets up data for NIMBLE -->
<!-- ## -->
<!-- ## Input: -->
<!-- ##    - functions/FXN-MVPv1.1.R -->
<!-- ##    - code/03-species-models/MVPv1.csv -->
<!-- ## -->
<!-- ## Output: -->
<!-- ##    - setup_BLOCK.rdata (saves environment for import to fit model) -->
<!-- ## -->
<!-- ## --------------------------- -->

So you wanna fit an integrated species distribution model? You've come to the right place!

# NEED SOME SORT OF INTRO HERE.

- The first part of this sets up the range, imports species and covariate data, scales covariates, and sets everything up for NIMBLE.
- The second part fits the NIMBLE model
- The third part summarizes the NIMBLE output and creates output figures.


Let's first load the libraries we'll need to set up, visualize, fit, and summarize our data.

```{r libraries, eval = F}
# Load libraries ----
library(tidyverse)
library(sf)
library(nimble)

## To install SpFut.flexiSDM or check for updates, use the commented code below:
# remotes::install_github("rileymummah/SpFut.flexiSDM", build_vignettes = T)
library(SpFut.flexiSDM)
```


## Model specifications

To load data or fit a model, you must first create a .csv file that outlines the model specifications. We call ours `model-specs.csv`. Let's load the .csv and take a look at how we structured it.

```{r model-specs}
mods <- read.csv("code/model-specs.csv")

head(mods, n=2)
```

The following table describes the column names, data type, and description found in `model-specs.csv`.

| Column Name     | Type      | Description                              |
|-----------------|-----------|------------------------------------------|
| number          | numeric   | Model number                             |
| sp.code         | character | Four digit species code                  |
| model           | character | Model name                               |
| mem.nim         | numeric   | Memory allocation (in MB) for NIMBLE. Could be omitted if not using an HPC. |
| mem.sum         | numeric   | Memory allocation (in MB) for parallelized summarization. Could be omitted if not using an HPC. |
| year.start      | numeric   | Starting year of data inclusion          |
| year.end        | numeric   | Ending year of data inclusion            |
| buffer          | numeric   | Size of buffer around range edge (in km) |
| sp.auto         | logical   | Should a spatial model be included?      |
| zero_mean       | logical   | Should a zero mean assumption be included in the spatial model |
| tau             | numeric/character | Assign a fixed value or a prior distribution in BUGS language for precision ($\tau$; e.g., dgamma(5,5)) |
| coarse.grid     | logical   | Should a coarse grid be used for the spatial model? Usually desired for large-range species |
| cont.grid       | logical   | Should the grid be restricted to continuous cells? (This will remove any groups of non-contiguous cells) |
| covs.PO         | character | A comma-separated list of covariates for modeling PO sampling effort. These must match the covariate names in the data. |
| covs.inat       | character | A comma-separated list of covariates for modeling iNaturalist sampling effort. These must match the covariate names in the data. |
| covs.lin        | character | A comma-separated list of linear covariates for modeling species relative abundance. These must match the covariate names in the data. |
| covs.quad       | character | A comma-separated list of quadratic covariates for modeling species relative abundance. These must match the covariate names in the data. |
| check.covs      | logical   | Should covariate selection remove multicollinearity? |
| Bprior          | character | Assign a prior distribution in BUGS lanaguage for the $\beta$ coefficients in the state model (e.g., dnorm(1,0))
| filter.region   | logical   | Should the region be spatially filtered?          |
| spat.bal        | logical   | Should the PO data be spatially balanced?         |
| coordunc        | numeric   | The level of acceptable coordinate uncertainty    |
| coordunc_na.rm  | logical   | Should coordinates with uncertainty beyond `coordunc` be removed? |
| block.rows      | numeric   | The number of rows for cross-validation blocks    |
| block.cols      | numeric   | The number of columns for cross-validation blocks |
| block.folds     | numeric   | The number of cross-validation block groups       |
| iter            | numeric   | The number of iterations to run an MCMC chain     |
| thin            | numeric   | Thin the MCMC chains by this parameter            |
| region.sub      | logical   | Only use the centroid of the region with a radius equal to the buffer size? |
| lon.hi          | numeric   | High longitude value to restrict the range
| lon.lo          | numeric   | Low longitude value to restrict the range
| lat.hi          | numeric   | High latitude value to restrict the range
| lat.lo          | numeric   | Low latitude value to restrict the range
| project         | numeric   | The number of future projections (otherwise, NA)  |
| exclude.dataset | character | A comma-separated list of datasets to exclude from analysis |


We have set up the scripts in the flexiSDM workflow so that a minimum number of parameters needs to be changed among them. To run this script in full (using our file arrangement), you only need to edit the following parameters:

```{r param.edit}
nums.do <- 3 # model number to run
block <- 'none' # block to run ('none', 1, 2, or 3)
local <- 1 # are you running the code locally (1) or on an HPC (0)?
```

For the purposes of this demonstration, we're going TO FIT X MODEL WITH THESE ASSUMPTIONS




### Load model specifications

Then we can use our model specifications to define a series of inputs for the setup script (found in `01-flexiSDM.R`).

```{r mods}
## Set up model variables
mods <- filter(mods, number %in% nums.do)

## Get variables for model from model-specs.csv
number <- mods$number[1] # model number
sp.code <- mods$sp.code[1] # 4-digit species code
model <- mods$model[1] # model name
sp.auto <- mods$sp.auto[1] # include spatial autocorrelation?
coarse.grid <- mods$coarse.grid[1] # use a coarse grid?
cont.grid <- mods$cont.grid[1] # restrict to only a continuous grid?
year.start <- mods$year.start[1] # start year for data
year.end <- mods$year.end[1] # end year for data
buffer <- mods$buffer[1] # buffer size (km)
filter.region <- mods$filter.region[1] # filter the region? LANE - FILTER BY WHAT?
spat.bal <- mods$spat.bal[1] # include spatial balancing?
coordunc <- mods$coordunc[1] # level of coordinate uncertainty to include (km)
coordunc_na.rm <- mods$coordunc_na.rm[1] # filter by coordinate uncertainty?
block.folds <- mods$block.folds[1] # how many groups of blocks?
block.rows <- mods$block.rows[1] # how many rows of blocks?
block.cols <- mods$block.cols[1] # how many columns of blocks?
block.out <- block # define the block to setup

if (block.out == "none") {
  blockname <- "full" # rename the block to 'full' if not doing cross-validation
} else {blockname <- block.out} # otherwise, keep the block number

# names of datasets to exclude (e.g., iNaturalist) - must match dataset naming
exclude.dataset <- unlist(str_split(mods$exclude.dataset[1], pattern = ", "))

## ICAR parameters
zero_mean <- mods$zero_mean[1] # zero mean assumption?
tau <- mods$tau[1] # precision (tau) can be a fixed value or a prior

## MCMC parameters
iter <- mods$iter[1] # number of MCMC iterations
thin <- mods$thin[1] # number to thin by
burnin <- floor(iter*0.75) # burnin to discard

## Change of region
region.sub <- mods$region.sub[1] # only estimate for a subregion?
lat.hi <- mods$lat.hi[1] # high value of latitude range
lat.lo <- mods$lat.lo[1] # low value of latitude range
lon.hi <- mods$lon.hi[1] # high value of longitude range
lon.lo <- mods$lon.lo[1] # low value of longitude range

## Future projections
project <- mods$project[1] # how many projections?
if (block.out != "none") {
  project <- 0 # no projections for cross-validation blocks
}

## Covariates
# list of PO covariates
covs.PO <- unlist(str_split(mods$covs.PO[1], pattern = ", "))
# list of iNaturalist covariates
covs.inat <- unlist(str_split(mods$covs.inat[1], pattern = ", "))
# list of linear covariates for state model
covs.lin <- unlist(str_split(mods$covs.lin[1], pattern = ", "))
# list of quadratic covariates for state model
covs.quad <- unlist(str_split(mods$covs.quad[1], pattern = ", "))
check.covs <- mods$check.covs[1] # covariate selection to remove multicollinearity?

# Define the prior for the state model coefficients
Bprior <- mods$Bprior[1]
```


Now that we have our model specifications loaded, we're ready to set up the region and load data and covariates. First, let's ensure that we have an output folder specific to the model we are fitting.

```{r output-folder}
# Make output folder ----
out.dir <- paste0("outputs/", number, "_", sp.code, "_", model, "/")
if (dir.exists(out.dir) == F) {
  dir.create(out.dir)
}
```

## Step 1: Define the region

We define the region of inference using the GAP and IUCN ranges. Other boundaries could be used to create the limits of the range. We saved the IUCN and GAP range boundaries in folders in `data/sp.code`, so that they can be called by the `get_range` function. We have also downloaded the US Census state boundary shapefile (`cb_2018_us_state_500k.shp`) to dictate the national and state borders. 

```{r ranges, results = 'hide'}
## Load ranges
range.path <- c(paste0("data/", sp.code, "/GAP/"),
                paste0("data/", sp.code, "/IUCN/"))
range.name <- c("GAP", "IUCN")
rangelist <- get_range(range.path,
                       range.name,
                       crs = 4326)

## Load USA boundary
exclude <- c("Alaska", "Hawaii", "Commonwealth of the Northern Mariana Islands",
             "American Samoa", "United States Virgin Islands", "Guam", "Puerto Rico")

usa <- st_read("data/USA/cb_2018_us_state_500k/cb_2018_us_state_500k.shp") %>%
        filter((NAME %in% exclude) == F) %>%
        st_union()

## Load CONUS grid covariates
load("data/USA/grid-covar.rdata")

## Generate the region for inference and plotting
region <- make_region(rangelist,
                      buffer = buffer,
                      crs = 3857,
                      sub = region.sub,
                      boundary = usa,
                      grid = conus.grid,
                      rm.clumps = T,
                      clump.size = 50,
                      continuous = cont.grid)

## If using the coarse spatial grid, redefine the grid and plot the resulting grid.
if (coarse.grid == T) {
  spatRegion <- suppressWarnings(make_spatkey(region$sp.grid))

  # Print spatial grid
  if (block.out == 'none') {
    pl <- ggplot(spatRegion$spat.grid) + geom_sf() + theme_bw()
    ggsave(pl, file = paste0(out.dir, "2_inputmap-e_spatGrid.jpg"), height = 8, width = 10)
  }
} else {
  spatRegion <- NULL
}
```



## Step 2: Set up cross-validation blocks (if using)

ADD SHORT EXPLANATION HERE

```{r cv-blocks}
## Set up cross validation blocks
spatblocks <- make_CV_blocks(region, rows = block.rows, cols = block.cols, k = block.folds)

## Set up training and test sets based on cross validation blocks
if (block.out == "none") { # If fitting the full model, then there are
  test.i <- c()            # no test data
  train.i <- region$sp.grid$conus.grid.id

} else {
  block1 <- spatblocks %>% filter(folds == block.out)

  ## Find grid.ids for test block, everything else is in the training block
  test.i <- st_intersection(region$sp.grid, block1) %>%
             pull(conus.grid.id) %>%
             unique()
  train.i <- filter(region$sp.grid, conus.grid.id %in% test.i == F) %>%
              pull(conus.grid.id)
}
```





LEAVE HERE WITH CONTEXT FOR CREATION AND USE
NEEDED FOR INDEXING IN NIMBLE CODE

```{r gridkey}
# Make gridkey ----
gridkey <- select(region$sp.grid, conus.grid.id) %>%
  st_drop_geometry() %>%
  mutate(group = case_when(conus.grid.id %in% train.i ~ "train",
                           conus.grid.id %in% test.i ~ "test"))
```





## Step 3: Load species data

TEXT AND CONTEXT HERE

```{r spp-codes}
## Define species codes
codeKey <- read.csv("data/model-specieslist.csv")

## Common name
common <- codeKey %>%
            filter(DS.code == sp.code) %>%
            pull(SSAR.common)
common

## All species codes used (some species are complexes or have subspecies so multiple codes exist)
sp.code.all <- codeKey %>%
                filter(DS.code == sp.code) %>%
                pull(all.codes)
sp.code.all

## Scientific name
sciname <- codeKey %>%
            filter(DS.code == sp.code) %>%
            pull(SSAR.scientific)
sciname

## Save genus name to assign Frog/Toad or Salamander
gen <- stringr::word(sciname, 1)

if (gen %in% c("Acris", "Anaxyrus", "Aquarana", "Ascaphus",
               "Craugastor", "Dendrobates", "Dryophytes", "Eleutherodactylus",
               "Gastrophryne", "Glandirana", "Hyla", "Hypopachus", "Incilius",
               "Leptodactylus", "Lithobates", "Osteopilus",
               "Pseudacris", "Rana", "Rhinella", "Rhinophrynus",
               "Scaphiopus", "Smilisca", "Spea", "Xenopus")) spp.type <- "Frog/Toad"
if (gen %in% c("Ambystoma", "Amphiuma", "Aneides", "Batrachoseps",
               "Cryptobranchus", "Desmognathus", "Dicamptodon",
               "Ensatina", "Eurycea", "Gyrinophilus", "Hemidactylium",
               "Hydromantes", "Necturus", "Notophthalmus",
               "Phaeognathus", "Plethodon", "Pseudobranchus",
               "Pseudotriton", "Rhyacotriton", "Siren",
               "Stereochilus", "Taricha", "Urspelerpes")) spp.type <- "Salamander"

gen; spp.type
```




<!-- # summarize datasets to get up-to-date available datasets -->
<!-- # this script also prints messages from QC functions -->
<!-- # source("DATA SWAMP/00-summarize-datasets.R") -- needed load_species_data() -->


```{r species-data, eval=F}
## Compile the name of all files that have data for the species of interest
allfiles <- read.csv("data/dataset-summary-full.csv")
# oNLY NEED VECTOR OF FILE NAMES AND FILE VECTORS
head(allfiles, 2)

tmp1 <- gsub("[|]", "$|^", sp.code.all) # THIS LINE WON'T RUN UNTIL THE SP.CODE.ALL IS CREATED
tmp2 <- paste0("^", tmp1, "$")
allfiles <- allfiles[grep(tmp2, allfiles$species),] %>%
            select(name, file) %>%
            distinct() %>%
            filter(name %in% exclude.dataset == F)
```


NEED THESE TO LOAD SPECIES DATA
NEED A LIST SAME LENGTH AS FILE VECTOR WHERE EACH ELEMENT IS A DETECTION COVARIATE

```{r load-species-data, eval=F}
# detection covariates for each of these datasets
covs <- read.csv("DATA SWAMP/00-data-summary-flexiSDM.csv") %>%
  filter(Data.Swamp.file.name %in% allfiles$file) %>%
  select(Name, Covar.mean, Covar.sum, Area)
covariates <- list()
for (i in 1:nrow(covs)) {
  covs.mean <- unlist(strsplit(covs$Covar.mean[i], split = ", "))
  covs.sum <- unlist(strsplit(covs$Covar.sum[i], split = ", "))
  #area <- unlist(strsplit(covs$Area[i], split = ","))
  covs1 <- c(covs.mean, covs.sum)
  covs1 <- covs1[which(is.na(covs1) == F)]
  covariates[[i]] <- covs1
}

species.data <- load_species_data(sp.code,
                                  sp.code.all,
                                  file.name = allfiles$file,
                                  file.label = allfiles$name,
                                  file.path = "DATA SWAMP/data-ready/",
                                  keep.cols = covariates,
                                  region = region,
                                  filter.region = filter.region,
                                  year.start = year.start,
                                  year.end = year.end,
                                  coordunc = coordunc,
                                  coordunc_na.rm = coordunc_na.rm,
                                  spat.thin = spat.bal,
                                  keep.conus.grid.id = gridkey$conus.grid.id[which(gridkey$group == "train")])
```




## Step 4: Plot species data


WE PROBABLY DON'T NEED TO INCLUDE ALL OF THE FIGURES THAT WE MAKE, SO LET'S CHOOSE A FEW TO HIGHLIGHT

```{r plot-species-data, eval=F}
## Quick title to be used across figures
if (block == "none") {
  title <- ", full model"
} else {
  title <- paste0(", excluding block ", block)
}

## Plot data samples
pl <- map_species_data(region = region,
                       species.data = species.data,
                       year.start = year.start,
                       year.end = year.end,
                       plot = "samples",
                       plot.region = T,
                       details = T,
                       title = paste0(common, " (", sp.code, ")", title))
## Save plot
# ggsave(pl, file = paste0(out.dir, "2_inputmap-b_data-", blockname, "-details.jpg"), height = 8, width = 10)
pl
```

<!-- pl <- map_species_data(region = region, -->
<!--                        species.data = species.data, -->
<!--                        year.start = year.start, -->
<!--                        year.end = year.end, -->
<!--                        plot = "samples", -->
<!--                        plot.region = T, -->
<!--                        details = F, -->
<!--                        title = paste0(common, " (", sp.code, ")", title)) -->
<!-- ggsave(pl, file = paste0(out.dir, "2_inputmap-a_data-", blockname, ".jpg"), height = 8, width = 10) -->


<!-- if (block.out != "none") { -->

<!--   pl <- map_species_data(region = region, -->
<!--                          species.data = species.data, -->
<!--                          year.start = year.start, -->
<!--                          year.end = year.end, -->
<!--                          plot = "samples", -->
<!--                          plot.blocks = T, -->
<!--                          blocks = spatblocks[which(spatblocks$folds == block),], -->
<!--                          plot.region = T, -->
<!--                          details = T, -->
<!--                          title = paste0(common, " (", sp.code, ")", title)) -->
<!--   ggsave(pl, file = paste0(out.dir, "2_inputmap-d_blocks-", blockname, "-details.jpg"), height = 8, width = 10) -->

<!--   pl <- map_species_data(region = region, -->
<!--                          species.data = species.data, -->
<!--                          year.start = year.start, -->
<!--                          year.end = year.end, -->
<!--                          plot = "samples", -->
<!--                          plot.blocks = T, -->
<!--                          blocks = spatblocks[which(spatblocks$folds == block),], -->
<!--                          plot.region = T, -->
<!--                          details = F, -->
<!--                          title = paste0(common, " (", sp.code, ")", title)) -->
<!--   ggsave(pl, file = paste0(out.dir, "2_inputmap-c_blocks-", blockname, ".jpg"), height = 8, width = 10) -->


<!-- } else { -->
<!--   pl <- map_species_data(region = region, -->
<!--                          species.data = species.data, -->
<!--                          year.start = year.start, -->
<!--                          year.end = year.end, -->
<!--                          plot = "samples", -->
<!--                          plot.blocks = T, -->
<!--                          blocks = spatblocks, -->
<!--                          plot.region = T, -->
<!--                          details = T, -->
<!--                          title = paste0(common, " (", sp.code, ")", title)) -->
<!--   ggsave(pl, file = paste0(out.dir, "2_inputmap-d_blocks-", blockname, "-details.jpg"), height = 8, width = 10) -->

<!--   pl <- map_species_data(region = region, -->
<!--                          species.data = species.data, -->
<!--                          year.start = year.start, -->
<!--                          year.end = year.end, -->
<!--                          plot = "samples", -->
<!--                          plot.blocks = T, -->
<!--                          blocks = spatblocks, -->
<!--                          plot.region = T, -->
<!--                          details = F, -->
<!--                          title = paste0(common, " (", sp.code, ")", title)) -->
<!--   ggsave(pl, file = paste0(out.dir, "2_inputmap-c_blocks-", blockname, ".jpg"), height = 8, width = 10) -->

<!-- } -->



## Load covariate data
<!-- load("data/USA/grid-covar.rdata") -->
<!-- covar <- conus.covar.grid %>% filter(conus.grid.id %in% region$sp.grid$conus.grid.id) -->
<!-- covar <- covar[order(match(covar$conus.grid.id, region$sp.grid$conus.grid.id)),] -->

<!-- # load iNat data for similar species if it exists -->
<!-- if ("iNaturalist" %in% names(species.data$obs)) { -->
<!--   cat("loading iNat records of supplemental species\n") -->

<!--   inat <- readRDS("data/USA/inateffortcov.rds") %>% -->
<!--     mutate(conus.grid.id = as.character(conus.grid.id)) -->

<!--   covar <- dplyr::left_join(covar, inat, by = c("conus.grid.id")) -->

<!--   # Assign Frog or Salamander iNat counts -->
<!--   if (spp.type == 'Frog/Toad') { -->
<!--     covar$n.inat <- covar$n.frog -->
<!--   } else { -->
<!--     covar$n.inat <- covar$n.sal -->
<!--   } -->

<!--   covar <- select(covar, -n.frog, -n.sal) -->


<!--   # Now check if records for this species need to be added -->

<!--   # read in all inat data for this species -->
<!--   dat <- read.csv(paste0("DATA SWAMP/data-ready/", gen, "_iNat_PO.csv")) %>% -->
<!--     filter(year >= 1994, year <= 2025) -->
<!--   tmp1 <- gsub("[|]", "$|^", sp.code.all) -->
<!--   tmp2 <- paste0("^", tmp1, "$") -->
<!--   dat <- dat[grep(tmp2, dat$species),] -->


<!--   # files <- list.files("../species-futures/DATA SWAMP/data-ready/", pattern = "_iNat_PO.csv") -->
<!--   # fs <- files[grep(sp.code.all, files)] -->
<!--   # if (length(fs) == 0) next -->
<!--   # inat.dat1 <- c() -->
<!--   # for (f in 1:length(fs)) { -->
<!--   #    -->
<!--   #   dat <- read.csv(paste0("../species-futures/DATA SWAMP/data-ready/", fs[f])) %>% -->
<!--   #     filter(year >= year.start, year <= year.end) -->
<!--   #    -->
<!--   #   # iNaturalist records for subspecies names might be duplicates, remove here -->
<!--   #   new <- paste0(dat$lat, dat$lon, dat$day, dat$month, dat$year) -->
<!--   #   old <- paste0(inat.dat1$lat, inat.dat1$lon, inat.dat1$day, inat.dat1$month, inat.dat1$year) -->
<!--   #    -->
<!--   #   if (length(old) > 0) { -->
<!--   #     rm <- which(new %in% old) -->
<!--   #     dat <- dat[-rm,] -->
<!--   #      -->
<!--   #     if (nrow(dat) == 0) { -->
<!--   #       #cat("All iNaturalist records are potential duplicates\n") -->
<!--   #       next -->
<!--   #     } else { -->
<!--   #       inat.dat1 <- bind_rows(inat.dat1, dat) -->
<!--   #     } -->
<!--   #   } else { -->
<!--   #     inat.dat1 <- bind_rows(inat.dat1, dat) -->
<!--   #   } -->
<!--   # } # end loading inat files -->


<!--   # if all points in a state are obscured (coord.unc > 25000), species was not included -->
<!--   tmp <- dat %>% -->
<!--     group_by(stateProvince) %>% -->
<!--     summarize(min.unc = min(coord.unc, na.rm = T)) %>% -->
<!--     mutate(min.unc = case_when(min.unc == Inf ~ NA, -->
<!--                                T ~ min.unc)) %>% -->
<!--     filter(min.unc > 25000) -->
<!--   obsc.state <- tmp$stateProvince -->
<!--   statecodes <- read.csv("data/statecodes.csv") -->
<!--   obsc.state <- statecodes$abbrev[which(statecodes$state %in% obsc.state)] -->


<!--   # species was already included -->
<!--   if (nrow(tmp) == 0) { -->


<!--     # Need to add species -->
<!--   } else { -->
<!--     cat("Adding", sp.code, "to iNat effort covariate\n") -->

<!--     inat.cells <- dat %>% -->
<!--       st_as_sf(coords = c("lon", "lat"), crs = 4326) %>% -->
<!--       st_transform(crs = 3857) %>% -->
<!--       st_intersection(region$sp.grid) %>% -->
<!--       st_drop_geometry() %>% -->
<!--       group_by(conus.grid.id) %>% -->
<!--       summarize(n.inat.sp = n()) %>% -->
<!--       full_join(region$sp.grid, ., by = "conus.grid.id") %>% -->
<!--       st_drop_geometry() -->
<!--     inat.cells$n.inat.sp[is.na(inat.cells$n.inat.sp)] <- 0 -->

<!--     covar <- full_join(covar, select(inat.cells, conus.grid.id, n.inat.sp), by = 'conus.grid.id') %>% -->
<!--       mutate(n.inat = n.inat + n.inat.sp) -->
<!--   } -->

<!-- } -->






<!-- # add centroid lat and lon of each grid cell to covar -->
<!-- centroid <- sf::st_centroid(region$sp.grid) %>% -->
<!--               sf::st_coordinates() %>% -->
<!--               as.data.frame() %>% -->
<!--               dplyr::mutate(conus.grid.id = region$sp.grid$conus.grid.id) -->

<!-- colnames(centroid)[1:2] <- c("lon", "lat") -->
<!-- covar <- dplyr::left_join(covar, centroid, by = "conus.grid.id") -->



<!-- ### Read in process covariates ---- -->
<!-- covs.z <- c(covs.lin, covs.quad, covs.int.factor) -->
<!-- if ("" %in% covs.z) { -->
<!--   covs.z <- covs.z[-which(covs.z == "")]   -->
<!-- } -->
<!-- if (NA %in% covs.z) { -->
<!--   covs.z <- covs.z[-which(is.na(covs.z))] -->
<!-- } -->



<!-- ### Add additional/derived covariates ----- -->
<!-- # If you add a new covariate, add a row to data/covariate-labels.csv with a label -->
<!-- if (sp.code == "ANMI") { # add in additional ANMI covariates -->

<!--   inv <- read_rds("data/species/ANMI/invsp-grid.rds") %>% -->
<!--     mutate(CRAY01 = case_when(CRAY_huc10 > 0 ~ 1, -->
<!--                               CRAY_huc10 == 0 ~ 0), -->
<!--            BULL01 = BULL_huc10, -->
<!--            FISH01 = case_when(FISH_huc10 > 0 ~ 1, -->
<!--                               FISH_huc10 == 0 ~ 0), -->
<!--            nonnative = CRAY01 + BULL01 + FISH01) -->
<!--   flow <- read_rds("data/species/ANMI/flow-grid.rds") -->
<!--   clade <- read_rds("data/species/ANMI/clade-grid.rds") -->
<!--   cwd <- read_rds("data/species/ANMI/cwd-grid.rds") -->
<!--   rip <- read_rds("data/species/ANMI/riparian-grid.rds") -->
<!--   rsvr <- read_rds("data/species/ANMI/rsvr-grid.rds") -->
<!--   aet <- read_rds("data/species/ANMI/aet-grid.rds") -->
<!--   perm <- read_rds("data/species/ANMI/waterperm-grid.rds") -->

<!--   covar <- left_join(covar, select(inv, !sp.grid.id), by = c("conus.grid.id")) %>% -->
<!--             left_join(select(flow, !sp.grid.id), by = c("conus.grid.id")) %>% -->
<!--             left_join(select(perm, !sp.grid.id), by = c("conus.grid.id")) %>% -->
<!--             left_join(select(aet, !sp.grid.id), by = c("conus.grid.id")) %>% -->
<!--             left_join(select(cwd, !sp.grid.id), by = c("conus.grid.id")) %>% -->
<!--             left_join(select(rip, !sp.grid.id), by = c("conus.grid.id")) %>% -->
<!--             left_join(select(rsvr, !sp.grid.id), by = c("conus.grid.id")) %>% -->
<!--             mutate(riparian = stream_forest) %>% -->
<!--             mutate(sqrt.mean.spring.discharge = sqrt(mean.spring.discharge), -->
<!--                    curt.mean.spring.discharge = mean.spring.discharge^(1/3)) %>% -->
<!--             left_join(select(ungroup(clade), !sp.grid.id), by = c("conus.grid.id")) -->


<!-- } else if (sp.code == "RACA") { -->

<!--   rain <- read_rds("data/species/RACA/rain-grid.rds") -->

<!--   covar <- left_join(covar, select(rain, !sp.grid.id), by = c("conus.grid.id")) %>% -->
<!--             mutate(sqrtarea_small = sqrt(area_small), -->
<!--                    sqrtarea_medium = sqrt(area_medium)) -->



<!-- } else if (sp.code == "GPOR") { -->

<!--   cwd <- read_rds("data/species/GPOR/cwd-grid.rds") -->
<!--   rain <- read_rds("data/species/GPOR/rain-grid.rds") -->
<!--   soilwater <- read_rds("data/species/GPOR/soilwater-grid.rds") -->
<!--   climate_velocity <- read_rds("data/species/GPOR/climate_velocity.rds") -->

<!--   covar <- left_join(covar, select(cwd, !sp.grid.id), by = c("conus.grid.id")) %>% -->
<!--             left_join(select(rain, !sp.grid.id), by = c("conus.grid.id")) %>% -->
<!--             left_join(select(soilwater, !sp.grid.id), by = c("conus.grid.id")) %>% -->
<!--             left_join(select(climate_velocity, !sp.grid.id), by = c("conus.grid.id")) -->


<!-- } else if (sp.code == "PLSE") { -->

<!--   if (buffer == 1) regs <- c("east", "west", "west", "west", "west") -->
<!--   if (buffer == 100000) regs <- c("east", "west", "west") -->
<!--   reg <- region$region %>% -->
<!--           st_cast("POLYGON") %>% -->
<!--           mutate(region = regs) -->

<!--   grid <- region$sp.grid %>% -->
<!--           st_intersection(reg) %>% -->
<!--           st_drop_geometry() -->

<!--   covar <- covar %>% -->
<!--             mutate(N_all = N + NW + NE) %>% -->
<!--             full_join(grid, by = c("conus.grid.id")) -->



<!-- } -->

<!-- rm <- which(complete.cases(covar[,covs.z]) == F) -->
<!-- if (length(rm) > 0) { -->
<!--   covar <- covar[-rm,] -->
<!--   region$sp.grid <- region$sp.grid[-rm,] -->
<!--   gridkey <- gridkey[-rm,] -->
<!-- } -->

<!-- # for 143_DFUS -->
<!-- if (number == 143) { -->

<!--   tmp <- species.data$locs$disc %>% -->
<!--     filter(source == "iNaturalist") %>% -->
<!--     left_join(covar, by = "conus.grid.id") -->

<!--   rm <- which(covar$footprint > 30) -->

<!--   if (length(rm) > 0) { -->


<!--     # first remove these from region, then find and remove other island cells -->
<!--     new.grid <- region$sp.grid[-rm,] -->
<!--     tmp <- new.grid %>% -->
<!--       # group neighboring cells -->
<!--       dplyr::summarize(geometry = sf::st_union(geometry, by_feature = F)) %>% -->
<!--       sf::st_cast("POLYGON")  -->

<!--     if (cont.grid == T) { -->
<!--       tmp1 <- tmp %>%  -->
<!--         mutate(area = as.numeric(sf::st_area(geometry))) %>% -->
<!--         arrange(desc(area)) %>%  -->
<!--         slice(1) -->
<!--     } else { -->
<!--       tmp1 <- tmp %>%  -->
<!--         mutate(area = as.numeric(sf::st_area(geometry))) %>% -->
<!--         dplyr::filter(area > 24999998 * 50) -->
<!--     } -->

<!--     tmp2 <- st_intersection(new.grid, tmp1)  -->
<!--     region$sp.grid <- tmp2 -->



<!--     covar <- filter(covar, conus.grid.id %in% region$sp.grid$conus.grid.id) -->
<!--     gridkey <- filter(gridkey, conus.grid.id %in% region$sp.grid$conus.grid.id) -->



<!--     # now put everything back in the correct order -->
<!--     covar <- covar[order(match(covar$conus.grid.id, region$sp.grid$conus.grid.id)),] -->
<!--     gridkey <- gridkey[order(match(gridkey$conus.grid.id, region$sp.grid$conus.grid.id)),] -->


<!--     # # fix spatRegion -->
<!--     # spatRegion$spatkey <- filter(spatRegion$spatkey, conus.grid.id %in% region$sp.grid$conus.grid.id) -->
<!--     # spatRegion$spat.grid <- filter(spatRegion$spat.grid, spat.grid.id %in% spatRegion$spatkey$spat.grid.id) -->
<!--     # spatRegion$spat.grid$spat.grid.id.new <- 1:nrow(spatRegion$spat.grid) -->
<!--     # spatRegion$spatkey <- full_join(spatRegion$spatkey, spatRegion$spat.grid, by = "spat.grid.id") -->

<!--   } -->



<!-- } -->





<!-- ### Scale covariates ---- -->
<!-- covar_unscaled <- covar -->

<!-- # scale numeric cols -->
<!-- numcols <- sapply(covar, is.numeric) -->
<!-- numcols <- which(numcols) -->
<!-- covar[,numcols] <- sapply(covar[,numcols], scale_this) -->

<!-- # remove covariates that are correlated > 0.4 -->
<!-- if (check.covs == T){ -->

<!--   threshold <- 0.4 -->
<!--   if (sp.code == "PJOR") threshold <- 0.45 -->

<!--   covs.rm <- select_covar(covs.z, threshold = threshold) -->
<!--   covs.lin <- covs.lin[-which(covs.lin %in% covs.rm)] -->
<!--   covs.quad <- covs.quad[-which(covs.quad %in% covs.rm)] -->

<!--   covs.z <- c(covs.lin, covs.quad) -->
<!-- } -->

<!-- if (length(covs.z) < 3) { -->
<!--   stop("There are fewer than 3 covariates remaining! This probably isn't a good model") -->
<!-- } -->


<!-- ### Clean up PO covariates ---- -->
<!-- if (length(covs.inat) == 1) { -->
<!--   if (covs.inat == "" | is.na(covs.inat)) covs.inat <- NULL -->
<!-- } -->
<!-- if (length(covs.PO) == 1) { -->
<!--   if (covs.PO == "" | is.na(covs.PO)) covs.PO <- NULL -->
<!-- } -->


<!-- ### Plot covariates ---- -->
<!-- if (block.out == "none") { -->


<!--   # Process covs -->

<!--   # get covariate labels -->
<!--   covlabs <- read.csv("data/covariate-labels.csv") %>% filter(covariate %in% covs.z) -->

<!--   plot_covar(covar, -->
<!--              region, -->
<!--              cov.names = covlabs$covariate, -->
<!--              cov.labels = covlabs$Label, -->
<!--              out.path = out.dir, -->
<!--              out.name = "1_covariates-a_process-map") -->

<!--   cor_covar(covar,  -->
<!--             cov.names = covlabs$covariate, -->
<!--             cov.labels = covlabs$Label, -->
<!--             out.path = out.dir, -->
<!--             out.name = "1_covariates-a_process-correlations",  -->
<!--             color.threshold = 0.25) -->



<!--   # iNat covs -->

<!--   # get covariate labels -->

<!--   covlabs <- read.csv("data/covariate-labels.csv") %>% -->
<!--     filter(covariate %in% covs.inat) -->

<!--   if (length(covs.inat) > 0) { -->
<!--     plot_covar(covar, -->
<!--                region, -->
<!--                cov.names = covlabs$covariate, -->
<!--                cov.labels = covlabs$Label, -->
<!--                out.path = out.dir, -->
<!--                out.name = "1_covariates-b_iNat-map") -->

<!--     if (length(covs.inat) > 1) { -->
<!--       cor_covar(covar,  -->
<!--                 cov.names = covlabs$covariate, -->
<!--                 cov.labels = covlabs$Label, -->
<!--                 out.path = out.dir, -->
<!--                 out.name = "1_covariates-b_iNat-correlations",  -->
<!--                 color.threshold = 0.25) -->
<!--     } -->
<!--   } -->



<!--   # PO covs -->

<!--   # get covariate labels -->
<!--   covlabs <- read.csv("data/covariate-labels.csv") %>% -->
<!--     filter(covariate %in% covs.PO) -->

<!--   if (length(covs.PO)) { -->
<!--     plot_covar(covar, -->
<!--                region, -->
<!--                cov.names = covlabs$covariate, -->
<!--                cov.labels = covlabs$Label, -->
<!--                out.path = out.dir, -->
<!--                out.name = "1_covariates-c_PO-map") -->

<!--     if (length(covs.PO) > 1) { -->
<!--       cor_covar(covar,  -->
<!--                 cov.names = covlabs$covariate, -->
<!--                 cov.labels = covlabs$Label, -->
<!--                 out.path = out.dir, -->
<!--                 out.name = "1_covariates-c_PO-correlations",  -->
<!--                 color.threshold = 0.25) -->
<!--     } -->
<!--   } -->

<!-- } -->



<!-- ### Add quadratic covariates and interactions ---- -->
<!-- if (length(covs.quad) > 0 & paste0(covs.quad, collapse = "") != "") { -->
<!--   for (c in 1:length(covs.quad)) { -->
<!--     covar[,paste0(covs.quad[c], "2")] <- covar[,covs.quad[c]] * covar[,covs.quad[c]] -->
<!--     covs.z <- c(covs.z, paste0(covs.quad[c], "2")) -->
<!--   } -->
<!-- } -->

<!-- # and interactions -->
<!-- if (length(covs.int.factor) == 1 & is.na(covs.int.factor) == F) { -->

<!--   # pivot factor wider to get dummy cols -->
<!--   covar[,covs.int.factor] <- paste0(covs.int.factor, ".", covar[,covs.int.factor]) -->
<!--   covar <- covar %>% -->
<!--             mutate(fact = 1) %>% -->
<!--             pivot_wider(names_from = all_of(covs.int.factor), values_from = fact, values_fill = 0) -->

<!--   covs.z <- c(covs.z, colnames(covar)[grep(paste0(covs.int.factor, "."), colnames(covar))]) -->

<!--   for (j in 1:length(covs.int.cont)) { -->
<!--     # add interaction -->
<!--     tmp <- add_int_cols(covar, int.factor = covs.int.factor, int.cont = covs.int.cont[j]) -->
<!--     covar <- tmp$covar -->
<!--     covs.z <- c(covs.z, tmp$names) -->
<!--   } -->

<!--   # remove reference level -->
<!--   rm <- grep(reference, covs.z) -->
<!--   covs.z <- covs.z[-rm] -->

<!--   # remove original column name -->
<!--   rm <- grep(covs.int.factor, covs.z)[1] -->
<!--   covs.z <- covs.z[-rm] -->
<!-- } -->




<!-- # Make spatkey ---- -->

<!-- if (coarse.grid == T) { -->
<!--   spatRegion <- suppressWarnings(make_spatkey(region$sp.grid)) -->

<!--   # Print spatial grid -->
<!--   if (block.out == 'none') { -->
<!--     pl <- ggplot(spatRegion$spat.grid) + geom_sf() + theme_bw() -->
<!--     ggsave(pl, file = paste0(out.dir, "2_inputmap-e_spatGrid.jpg"), height = 8, width = 10) -->
<!--   } -->
<!-- } else { -->
<!--   spatRegion <- NULL -->
<!-- } -->






<!-- # NIMBLE ---- -->

<!-- # add grid.id to gridkey -->
<!-- gridkey$grid.id <- 1:nrow(gridkey) -->




<!-- summary <- read.csv("DATA SWAMP/00-data-summary-flexiSDM.csv") %>% -->
<!--             filter(Data.Swamp.file.name %in% allfiles$file, -->
<!--                    Name %in% names(species.data$obs)) %>% -->
<!--   select(-Data.Swamp.file.name) %>% -->
<!--   distinct() -->
<!-- summary <- summary[order(match(summary$Name, names(species.data$obs))),] %>% distinct() -->

<!-- sp.data <- sppdata_for_nimble(species.data, -->
<!--                               region, -->
<!--                               data.type = summary$Type.true, -->
<!--                               PO.extent = summary$PO.extent, -->
<!--                               covar = covar, -->
<!--                               covs.inat = covs.inat, -->
<!--                               covs.PO = covs.PO, -->
<!--                               covs.mean = summary$Covar.mean, -->
<!--                               covs.sum = summary$Covar.sum, -->
<!--                               offset.area = rep("", nrow(summary)), -->
<!--                               DND.maybe = 1, -->
<!--                               keep.conus.grid.id = gridkey$conus.grid.id[which(gridkey$group == "train")]) # get rid of PO cells that are in the wrong grid cells -->


<!-- ### Data/constants ---- -->
<!-- tmp <- data_for_nimble(sp.data, covar = covar, covs.z, -->
<!--                        sp.auto = sp.auto, coarse.grid = coarse.grid, region = region, -->
<!--                        process.intercept = process.intercept, -->
<!--                        gridkey = gridkey, spatRegion= spatRegion) -->

<!-- data <- tmp$data -->
<!-- constants <- tmp$constants -->


<!-- # Add state indicator variable for iNat data to indicate which states have taxon geoprivacy -->
<!-- # Add state indicator for multi-state PO to indicate which states have data -->
<!-- constants <- add_state_ind(species.data, -->
<!--                            region, -->
<!--                            gridkey, -->
<!--                            constants, -->
<!--                            covs.inat = covs.inat, -->
<!--                            obsc.state = obsc.state, -->
<!--                            keep.conus.grid.id = gridkey$conus.grid.id[which(gridkey$group == "train")]) -->


<!-- if (sp.code == "ANMI") { -->

<!--   # Need to replace NMGFD datasets with Bernoulli instead of occupancy -->
<!--   for (s in 1:length(species.data$obs)) { -->
<!--     nm <- constants[[paste0("name", s)]] -->
<!--     nm <- gsub(" .*", "", nm) -->

<!--     if (nm %in% c("NMGFD", "AZGFD")) { -->
<!--       constants[[paste0("nVisitsV", s)]] <- 1 -->
<!--     } -->
<!--   } -->

<!-- } -->



<!-- if (model == "proj08nostate") { -->
<!--   if (sp.code == "EBIS") { -->
<!--     data$Xw3$WV <- NULL -->
<!--     data$Xw3$VT <- NULL -->
<!--     data$S3 <- NULL -->

<!--     constants$nCovW3 <- ncol(data$Xw3) -->
<!--   } else { -->
<!--     stop("Don't know which states to remove") -->
<!--   } -->

<!--   rm.state <- T -->
<!-- } else { -->
<!--   rm.state <- F -->
<!-- } -->



<!-- ### Code ---- -->
<!-- code <- nimble_code(data, -->
<!--                     constants,  -->
<!--                     path = out.dir, -->
<!--                     sp.auto = sp.auto,  -->
<!--                     coarse.grid = coarse.grid, -->
<!--                     Bpriordist = Bpriordist, Bpriorvar1 = Bpriorvar1, Bpriorvar2 = Bpriorvar2, -->
<!--                     block.out = block.out, -->
<!--                     min.visits.incl = 3,  -->
<!--                     zero_mean = zero_mean, -->
<!--                     tau = tau, -->
<!--                     rm.state = rm.state) -->

<!-- ### Initial values ---- -->
<!-- inits <- function(x){nimble_inits(data, -->
<!--                                   constants, -->
<!--                                   sp.auto = sp.auto, -->
<!--                                   seed = x)} -->

<!-- ### Parameters ---- -->
<!-- params <- nimble_params(data, -->
<!--                         constants, -->
<!--                         lambda = T, -->
<!--                         XB = T, -->
<!--                         sp.auto = sp.auto, -->
<!--                         effort = T) -->


<!-- source("code/03-species-models/xx-flexiSDM-setuptests.R") -->


<!-- end1 <- Sys.time() - start1 -->

<!-- # Remove local and block in case the setup is run locally but the model is fit on the HPC. -->
<!-- # Remove other unnecessary files to reduce the size of setup_BLOCK.Rdata -->
<!-- rm(list=c('local','block','args','conus.covar.grid','conus.grid','usa','conus', -->
<!--           'pl','centroid')) -->


<!-- # Save environment and full set up -->
<!-- save.image(paste0(out.dir, "setup_",block.out,".Rdata")) -->


<!-- # End script - proceed to 02-flexiSDM.R -->

<!-- ## --------------------------- -->
<!-- ## Objective:  -->
<!-- ##    - Fit NIMBLE model using setup_BLOCK.rdata file -->
<!-- ##  -->
<!-- ## Input: -->
<!-- ##    - setup_BLOCK.rdata -->
<!-- ## -->
<!-- ## Output:  -->
<!-- ##    - samples_BLOCK_CHAIN.rds OR samples_BLOCK.rds -->
<!-- ## -->
<!-- ## --------------------------- -->

<!-- start2 <- Sys.time()  -->
<!-- print(paste0('Beginning 02-MVPv1.1 script at ', start2)) -->


<!-- # EDIT THIS SECTION ---- -->
<!-- num <- 1 -->
<!-- block <- "none" -->
<!-- sp.code <- "RACA" -->
<!-- model <- "WIPtest" -->
<!-- chain <- 1 -->
<!-- local <- 1 -->
<!-- # --- -->


<!-- args = commandArgs(trailingOnly = TRUE) -->

<!-- if (length(args) > 0) { -->
<!--   num = as.numeric(args[1]) -->
<!--   sp.code = as.character(args[2]) -->
<!--   model = as.character(args[3]) -->
<!--   block = as.numeric(args[4]) -->
<!--   chain = as.numeric(args[5]) -->
<!--   local = as.numeric(args[6])  -->

<!--   # If running on HPC, set working directory -->
<!--   if (local == 0) { -->
<!--     setwd('/caldera/hovenweep/projects/usgs/ecosystems/eesc/rmummah/species-futures/') -->
<!--   }  -->
<!-- } -->

<!-- # Ensure block is coded correctly -->
<!-- if (block == 4) { -->
<!--   block <- 'none' -->
<!-- } -->


<!-- # Load functions and packages -->
<!-- suppressMessages(source("functions/FXN-nimbleParallel.R")) -->
<!-- library(SpFut.flexiSDM) -->


<!-- # Set output directory and load setup file -->
<!-- out.dir = paste0('outputs/03-species-models/MVPv1/',num,'_',sp.code,'_',model,'/') -->
<!-- load(paste0(out.dir,'setup_',block,'.Rdata')) -->


<!-- # Fit NIMBLE model ---- -->
<!-- if (local == 1) { -->
<!--   start.nim <- Sys.time() -->
<!--   samples <- nimbleParallel(code = code, -->
<!--                             data = data, -->
<!--                             constants = constants, -->
<!--                             inits = inits, -->
<!--                             param = params, -->
<!--                             iter = iter, -->
<!--                             burnin = burnin, -->
<!--                             thin = thin) -->

<!--   end.nim <- Sys.time() - start.nim -->
<!--   print(end.nim) -->

<!--   saveRDS(samples, paste0(out.dir,'samples_',block,'.rds')) -->
<!-- } else { -->

<!--   info <- list(seed = chain, -->
<!--                inits = inits(chain)) -->

<!--   start.nim <- Sys.time() -->
<!--   samples <- run_nimbleMCMC(info = info,  -->
<!--                             code = code,  -->
<!--                             constants = constants,  -->
<!--                             data = data,  -->
<!--                             param = params,  -->
<!--                             iter = iter,  -->
<!--                             burnin = burnin,  -->
<!--                             thin = thin) -->

<!--   end.nim <- Sys.time() - start.nim -->
<!--   print(end.nim) -->

<!--   saveRDS(samples, paste0(out.dir,'samples_',block,'_',chain,'.rds')) -->
<!-- } -->


<!-- end2 <- Sys.time() - start2 -->
<!-- saveRDS(end2, paste0(out.dir,'time2_',block,'.rds')) -->
<!-- print(end2) -->

<!-- # End script - proceed to 03-flexiSDM.R -->

<!-- ## --------------------------- -->
<!-- ## Objective:  -->
<!-- ##    - Fit NIMBLE model using setup_BLOCK.rdata file -->
<!-- ##  -->
<!-- ## Input: -->
<!-- ##    - setup_BLOCK.rdata -->
<!-- ##    - samples_BLOCK_CHAIN.rds OR samples_BLOCK.rds -->
<!-- ## -->
<!-- ## Output:  -->
<!-- ##    - samples_BLOCK_CHAIN.rds OR samples_BLOCK.rds -->
<!-- ## -->
<!-- ## --------------------------- -->

<!-- start3 <- Sys.time() -->


<!-- ## load packages ---- -->
<!-- library(tidyverse, quietly = T) -->
<!-- library(magrittr, quietly = T) -->
<!-- library(sf, quietly = T) -->
<!-- library(SpFut.flexiSDM) -->



<!-- # EDIT THIS SECTION ---- -->
<!-- num <- 108 -->
<!-- block <- "none" -->
<!-- sp.code <- "RACA" -->
<!-- model <- "tau1" -->
<!-- maxchain <- 3 -->
<!-- local <- 1 -->
<!-- # --- -->


<!-- # Setup ---- -->
<!-- args = commandArgs(trailingOnly = TRUE) -->

<!-- if (length(args) > 0) { -->
<!--   num = as.numeric(args[1]) -->
<!--   sp.code = as.character(args[2]) -->
<!--   model = as.character(args[3]) -->
<!--   block = as.numeric(args[4]) -->
<!--   maxchain = as.numeric(args[5]) -->
<!--   local = as.numeric(args[6]) -->

<!--   if (local == 0) { -->
<!--     setwd('/caldera/hovenweep/projects/usgs/ecosystems/eesc/rmummah/species-futures/') -->
<!--   }  -->

<!-- } -->

<!-- # Ensure block is coded correctly -->
<!-- if (block == 4) { -->
<!--   block <- 'none' -->
<!-- } -->

<!-- # load output directory, setup, and functions -->
<!-- out.dir = paste0('outputs/03-species-models/MVPv1/',num,'_',sp.code,'_',model,'/') -->
<!-- load(paste0(out.dir,'setup_',block,'.Rdata')) -->




<!-- # Load chains (samples) ---- -->

<!-- if (file.exists(paste0(out.dir,'samples_',block,'.rds'))) { # If it was run locally, this file should exist -->
<!--   samples <- readRDS(paste0(out.dir,'samples_',block,'.rds')) -->
<!-- } else { -->
<!--   # Load individual sample files and save as a list -->
<!--   samples <- lapply(1:maxchain,  -->
<!--                     FUN = function(x) { -->
<!--                       readRDS(paste0(out.dir,'samples_',block,'_',x,'.rds')) -->
<!--                     }) -->
<!-- } -->



<!-- ## Calculate derived quantities ---- -->
<!-- # This takes longer with the coarse grid. -->
<!-- samples <- lapply(samples, get_derived, data = data, project = project,  -->
<!--                   coarse.grid = coarse.grid, spatRegion = spatRegion, -->
<!--                   pathToProj = 'code/03-species-models/setup-projections.R') -->


<!-- # Plot chains before summarizing -->
<!-- cov.labs <- read.csv("data/covariate-labels.csv") -->

<!-- ### Chains ---- -->
<!-- ggsave(plot_chains(samples, data = data, cov.labs = cov.labs, plot = "B", cutoff = 0), file = paste0(out.dir, "3_parameters-a1_chains-B-",block,".jpg"), height = 6, width = 8) -->
<!-- ggsave(plot_chains(samples, data = data, cov.labs = cov.labs, plot = "alpha", cutoff = 0), file = paste0(out.dir, "3_parameters-b1_chains-alpha-",block,".jpg"), height = 6, width = 8) -->
<!-- ggsave(plot_chains(samples, data = data, cov.labs = cov.labs, plot = "tau", cutoff = 0), file = paste0(out.dir, "3_parameters-b1_chains-tau-",block,".jpg"), height = 6, width = 8) -->

<!-- # Posteriors ---- -->
<!-- ggsave(plot_posteriors(samples, data = data, cov.labs = cov.labs, plot = "B", cutoff = 0), file = paste0(out.dir, "3_parameters-a2_posteriors-B.jpg"), height = 6, width = 8) -->
<!-- ggsave(plot_posteriors(samples, data = data, cov.labs = cov.labs, plot = "alpha", cutoff = 0), file = paste0(out.dir, "3_parameters-b2_posteriors-alpha.jpg"), height = 6, width = 8) -->


<!-- ## Summarize chains ---- -->
<!-- out <- summarize_samples(samples,  -->
<!--                          data,  -->
<!--                          constants,  -->
<!--                          project = project, -->
<!--                          coarse.grid = coarse.grid, -->
<!--                          block.out = block.out, -->
<!--                          gridkey = gridkey,  -->
<!--                          spatkey = spatRegion$spatkey, -->
<!--                          effort = T, -->
<!--                          SLURM = ifelse(local == 1, F, T)) -->

<!-- save(out, file = paste0(out.dir, "data", blockname, ".rdata")) -->



<!-- # Plot output ---- -->

<!-- # Convergence -->
<!-- ggsave(plot_convergence(out), file = paste0(out.dir, "3_parameters-0_convergence-", block, ".jpg"), -->
<!--        height = 5, width = 12) -->

<!-- if (block.out == "none") { -->

<!--   ### Parameter estimates ---- -->
<!--   ggsave(plot_pars(out,  -->
<!--                    plot.type = "full",  -->
<!--                    plot.group = "process",  -->
<!--                    title = "Process parameter estimates", -->
<!--                    cov.labs = cov.labs), -->
<!--          file = paste0(out.dir, "3_parameters-a3_B.jpg"), height = 6, width = 10) -->

<!--   ggsave(plot_pars(out,  -->
<!--                    plot.type = "full",  -->
<!--                    plot.group = "observation",  -->
<!--                    title = "Observation parameter estimates", -->
<!--                    cov.labs = cov.labs), -->
<!--          file = paste0(out.dir, "3_parameters-b3_alpha.jpg"), height = 6, width = 10) -->

<!--   ggsave(plot_pars(out,  -->
<!--                    plot.type = "full",  -->
<!--                    plot.group = "dataset",  -->
<!--                    title = "Dataset intercept estimates"), -->
<!--          file = paste0(out.dir, "3_parameters-c3_observation.jpg"), height = 6, width = 10) -->

<!--   ggsave(plot_pars(out,  -->
<!--                    plot.type = "full",  -->
<!--                    plot.group = "tau",  -->
<!--                    title = "Tau estimate"), -->
<!--          file = paste0(out.dir, "3_parameters-c3_tau.jpg"), height = 6, width = 10) -->

<!--   ### Marginal effects ---- -->
<!--   ggsave(plot_effects(data, out, breaks = 0.001), file = paste0(out.dir, "/3_parameters-a4-effects.jpg"), height = 7, width = 10) -->




<!--   ### Plot effort and PO records ---- -->

<!--   tmp <- full_join(select(out$effort, conus.grid.id, PO.dataset.name, mean), select(out$lambda0, conus.grid.id, mean), by = "conus.grid.id") %>% -->
<!--     rename(mean.effort = mean.x, -->
<!--            mean.lambda = mean.y) -->

<!--   ind <- grep("nW", names(constants)) -->
<!--   nums <- gsub("nW", "", names(constants)[ind]) -->

<!--   dat <- c() -->
<!--   for (i in 1:length(nums)) { -->
<!--     dat1 <- data.frame(W = data[[paste0("W", nums[i])]], -->
<!--                        grid.id = constants[[paste0("Wcells", nums[i])]], -->
<!--                        PO.dataset.name = constants[[paste0("name", nums[i])]]) -->

<!--     dat <- bind_rows(dat, dat1) -->
<!--   } -->
<!--   dat <- dat %>% -->
<!--     full_join(gridkey, by = "grid.id") -->

<!--   tmp1 <- full_join(tmp, dat, by = c("conus.grid.id", "PO.dataset.name")) %>% -->
<!--     mutate(recpereff = W / mean.effort, -->
<!--            adjlambda = mean.effort * mean.lambda) -->

<!--   pl <- ggplot(tmp1) + -->
<!--     geom_point(aes(x = recpereff, y = mean.lambda), alpha = 0.5) + -->
<!--     geom_smooth(aes(x = recpereff, y = mean.lambda), method = "lm") + -->
<!--     facet_wrap(~ PO.dataset.name, scales = "free") + -->
<!--     theme_bw() + -->
<!--     labs(x = "Number of records per unit of effort", y = "Estimated lambda") -->
<!--   ggsave(pl, file = paste0(out.dir, "/X_recordspereffort.jpg"), height = 8, width = 8) -->

<!--   pl <- ggplot(tmp1) + -->
<!--     geom_point(aes(x = mean.effort, y = W), alpha = 0.5) + -->
<!--     geom_smooth(aes(x = mean.effort, y = W), method = "lm") + -->
<!--     facet_wrap(~ PO.dataset.name, scales = "free") + -->
<!--     theme_bw() + -->
<!--     labs(x = "Estimated effort", y = "Number of records") -->
<!--   ggsave(pl, file = paste0(out.dir, "/X_recordsvseffort.jpg"), height = 8, width = 8) -->


<!--   ### Maps ---- -->
<!--   # Current -->
<!--   pl <- map_species_data(title = paste0(common, " (", sp.code, ")"),  -->
<!--                          region,  -->
<!--                          plot = "lambda", -->
<!--                          out = out) -->
<!--   ggsave(pl, file = paste0(out.dir, "4_mapcurrent-a_lambda.jpg"), height = 7, width = 9) -->

<!--   pl <- map_species_data(title = paste0(common, " (", sp.code, ")"), -->
<!--                          region,  -->
<!--                          plot = "lambda", -->
<!--                          out = out,  -->
<!--                          plot.uncertainty = "unc.range") -->
<!--   ggsave(pl, file = paste0(out.dir, "4_mapcurrent-a_lambda-uncabs.jpg"), height = 7, width = 9) -->

<!--   pl <- map_species_data(title = paste0(common, " (", sp.code, ")"), -->
<!--                          region, -->
<!--                          plot = "lambda", -->
<!--                          out = out,  -->
<!--                          plot.uncertainty = "unc.rel") -->
<!--   ggsave(pl, file = paste0(out.dir, "4_mapcurrent-a_lambda-uncrel.jpg"), height = 7, width = 9) -->

<!--   pl <- map_species_data(title = paste0(common, " (", sp.code, ")"), -->
<!--                          region, -->
<!--                          plot = "psi", -->
<!--                          out = out) -->
<!--   ggsave(pl, file = paste0(out.dir, "4_mapcurrent-b_psi.jpg"), height = 7, width = 9) -->

<!--   pl <- map_species_data(title = paste0(common, " (", sp.code, ")"), -->
<!--                          region, -->
<!--                          plot = "psi", -->
<!--                          out = out,  -->
<!--                          plot.uncertainty = "unc.range") -->
<!--   ggsave(pl, file = paste0(out.dir, "4_mapcurrent-b_psi-uncabs.jpg"), height = 7, width = 9) -->

<!--   pl <- map_species_data(title = paste0(common, " (", sp.code, ")"), -->
<!--                          region, -->
<!--                          plot = "psi", -->
<!--                          out = out,  -->
<!--                          plot.uncertainty = "unc.rel") -->
<!--   ggsave(pl, file = paste0(out.dir, "4_mapcurrent-b_psi-uncrel.jpg"), height = 7, width = 9) -->

<!--   pl <- map_species_data(title = paste0(common, " (", sp.code, ")"), -->
<!--                          region, -->
<!--                          plot = "boundary", -->
<!--                          out = out,  -->
<!--                          threshold = 0.25) -->
<!--   ggsave(pl, file = paste0(out.dir, "4_mapcurrent-c1_boundary-0.25.jpg"), height = 7, width = 9) -->

<!--   pl <- map_species_data(title = paste0(common, " (", sp.code, ")"), -->
<!--                          region, -->
<!--                          plot = "boundary", -->
<!--                          out = out,  -->
<!--                          threshold = 0.5) -->
<!--   ggsave(pl, file = paste0(out.dir, "4_mapcurrent-c2_boundary-0.5.jpg"), height = 7, width = 9) -->

<!--   pl <- map_species_data(title = paste0(common, " (", sp.code, ")"), -->
<!--                          region, -->
<!--                          plot = "boundary", -->
<!--                          out = out,  -->
<!--                          threshold = 0.75) -->
<!--   ggsave(pl, file = paste0(out.dir, "4_mapcurrent-c3_boundary-0.75.jpg"), height = 7, width = 9) -->

<!--   if (sp.auto == T) { -->
<!--     pl <- map_species_data(title = paste0(common, " (", sp.code, ")"), -->
<!--                            region, -->
<!--                            plot = "spat", -->
<!--                            out = out, -->
<!--                            coarse.grid = coarse.grid, -->
<!--                            spatRegion = spatRegion) -->
<!--     ggsave(pl, file = paste0(out.dir, "4_mapcurrent-d_spat.jpg"), height = 7, width = 9) -->
<!--   } -->

<!--   pl <- map_species_data(title = paste0(common, " (", sp.code, ")"), -->
<!--                          region, -->
<!--                          plot = "XB", -->
<!--                          out = out, -->
<!--                          plot.exp = T) -->
<!--   ggsave(pl, file = paste0(out.dir, "4_mapcurrent-d_expXB.jpg"), height = 7, width = 9) -->


<!--   pl <- map_species_data(title = paste0(common, " (", sp.code, ")"),  -->
<!--                          region,  -->
<!--                          plot = "effort", -->
<!--                          out = out) -->
<!--   ggsave(pl, file = paste0(out.dir, "4_mapcurrent-e_effort.jpg"), height = 7, width = 9) -->

<!--   ### Future projections ---- -->
<!--   if (project > 0) { -->
<!--     pl <- map_species_data(title = paste0(common, " (", sp.code, ")"), -->
<!--                            region, -->
<!--                            plot = "lambda", -->
<!--                            out = out,  -->
<!--                            plot.current = F,  -->
<!--                            plot.change = F, -->
<!--                            proj.names = proj.names) -->
<!--     ggsave(pl, file = paste0(out.dir, "5_mapfuture-a_lambda.jpg"), height = 7, width = 9) -->

<!--     pl <- map_species_data(title = paste0(common, " (", sp.code, ")"), -->
<!--                            region, -->
<!--                            plot = "lambda", -->
<!--                            out = out,  -->
<!--                            plot.current = F,  -->
<!--                            plot.change = F,  -->
<!--                            plot.uncertainty = "unc.range", -->
<!--                            proj.names = proj.names) -->
<!--     ggsave(pl, file = paste0(out.dir, "5_mapfuture-a_lambda-uncabs.jpg"), height = 7, width = 9) -->

<!--     pl <- map_species_data(title = paste0(common, " (", sp.code, ")"), -->
<!--                            region, -->
<!--                            plot = "lambda", -->
<!--                            out = out,  -->
<!--                            plot.current = F,  -->
<!--                            plot.change = F,  -->
<!--                            plot.uncertainty = "unc.rel", -->
<!--                            proj.names = proj.names) -->
<!--     ggsave(pl, file = paste0(out.dir, "5_mapfuture-a_lambda-uncrel.jpg"), height = 7, width = 9) -->

<!--     pl <- map_species_data(title = paste0(common, " (", sp.code, ")"), -->
<!--                            region,  -->
<!--                            plot = "psi", -->
<!--                            out = out,  -->
<!--                            plot.current = F,  -->
<!--                            plot.change = F, -->
<!--                            proj.names = proj.names) -->
<!--     ggsave(pl, file = paste0(out.dir, "5_mapfuture-b_psi.jpg"), height = 7, width = 9) -->

<!--     pl <- map_species_data(title = paste0(common, " (", sp.code, ")"), -->
<!--                            region, -->
<!--                            plot = "psi", -->
<!--                            out = out,  -->
<!--                            plot.current = F,  -->
<!--                            plot.change = F,  -->
<!--                            plot.uncertainty = "unc.range", -->
<!--                            proj.names = proj.names) -->
<!--     ggsave(pl, file = paste0(out.dir, "5_mapfuture-b_psi-uncabs.jpg"), height = 7, width = 9) -->

<!--     pl <- map_species_data(title = paste0(common, " (", sp.code, ")"), -->
<!--                            region, -->
<!--                            plot = "psi", -->
<!--                            out = out,  -->
<!--                            plot.current = F,  -->
<!--                            plot.change = F,  -->
<!--                            plot.uncertainty = "unc.rel", -->
<!--                            proj.names = proj.names) -->
<!--     ggsave(pl, file = paste0(out.dir, "5_mapfuture-b_psi-uncrel.jpg"), height = 7, width = 9) -->

<!--     pl <- map_species_data(title = paste0(common, " (", sp.code, ")"), -->
<!--                            region, -->
<!--                            plot = "boundary", -->
<!--                            out = out,  -->
<!--                            plot.current = F,  -->
<!--                            plot.change = F,  -->
<!--                            threshold = 0.25, -->
<!--                            proj.names = proj.names) -->
<!--     ggsave(pl, file = paste0(out.dir, "5_mapfuture-c1_boundary-0.25.jpg"), height = 7, width = 9) -->

<!--     pl <- map_species_data(title = paste0(common, " (", sp.code, ")"), -->
<!--                            region, -->
<!--                            plot = "boundary", -->
<!--                            out = out,  -->
<!--                            plot.current = F,  -->
<!--                            plot.change = F,  -->
<!--                            threshold = 0.5, -->
<!--                            proj.names = proj.names) -->
<!--     ggsave(pl, file = paste0(out.dir, "5_mapfuture-c2_boundary-0.5.jpg"), height = 7, width = 9) -->

<!--     pl <- map_species_data(title = paste0(common, " (", sp.code, ")"), -->
<!--                            region, -->
<!--                            plot = "boundary", -->
<!--                            out = out,  -->
<!--                            plot.current = F,  -->
<!--                            plot.change = F,  -->
<!--                            threshold = 0.75, -->
<!--                            proj.names = proj.names) -->
<!--     ggsave(pl, file = paste0(out.dir, "5_mapfuture-c3_boundary-0.75.jpg"), height = 7, width = 9) -->

<!--     pl <- map_species_data(title = paste0(common, " (", sp.code, ")"), -->
<!--                            region, -->
<!--                            plot = "XB", -->
<!--                            out = out,  -->
<!--                            plot.exp = T, -->
<!--                            plot.current = F, -->
<!--                            proj.names = proj.names) -->
<!--     ggsave(pl, file = paste0(out.dir, "5_mapfuture-d_expXB-future.jpg"), height = 7, width = 9) -->


<!--     ### Future absolute change ---- -->
<!--     pl <- map_species_data(title = paste0(common, " (", sp.code, ")"), -->
<!--                            region, -->
<!--                            plot = "lambda", -->
<!--                            out = out,  -->
<!--                            plot.current = F,  -->
<!--                            plot.change = "absolute", -->
<!--                            proj.names = proj.names) -->
<!--     ggsave(pl, file = paste0(out.dir, "6_mapchange-a1_lambda-abs.jpg"), height = 7, width = 9) -->

<!--     pl <- map_species_data(title = paste0(common, " (", sp.code, ")"), -->
<!--                            region, -->
<!--                            plot = "psi", -->
<!--                            out = out,  -->
<!--                            plot.current = F,  -->
<!--                            plot.change = "absolute", -->
<!--                            proj.names = proj.names) -->
<!--     ggsave(pl, file = paste0(out.dir, "6_mapchange-b1_psi-abs.jpg"), height = 7, width = 9) -->

<!--     pl <- map_species_data(title = paste0(common, " (", sp.code, ")"),  -->
<!--                            region, -->
<!--                            plot = "boundary", -->
<!--                            out = out,  -->
<!--                            plot.current = F,  -->
<!--                            plot.change = "absolute",  -->
<!--                            threshold = 0.25, -->
<!--                            proj.names = proj.names) -->
<!--     ggsave(pl, file = paste0(out.dir, "6_mapchange-c1_boundary-abs-0.25.jpg"), height = 7, width = 9) -->

<!--     pl <- map_species_data(title = paste0(common, " (", sp.code, ")"),  -->
<!--                            region, -->
<!--                            plot = "boundary", -->
<!--                            out = out,  -->
<!--                            plot.current = F,  -->
<!--                            plot.change = "absolute",  -->
<!--                            threshold = 0.5, -->
<!--                            proj.names = proj.names) -->
<!--     ggsave(pl, file = paste0(out.dir, "6_mapchange-c2_boundary-abs-0.5.jpg"), height = 7, width = 9) -->

<!--     pl <- map_species_data(title = paste0(common, " (", sp.code, ")"),  -->
<!--                            region, -->
<!--                            plot = "boundary", -->
<!--                            out = out,  -->
<!--                            plot.current = F,  -->
<!--                            plot.change = "absolute",  -->
<!--                            threshold = 0.75, -->
<!--                            proj.names = proj.names) -->
<!--     ggsave(pl, file = paste0(out.dir, "6_mapchange-c3_boundary-abs-0.75.jpg"), height = 7, width = 9) -->

<!--     ### Future relative change ---- -->
<!--     pl <- map_species_data(title = paste0(common, " (", sp.code, ")"),  -->
<!--                            region, -->
<!--                            plot = "lambda", -->
<!--                            out = out,  -->
<!--                            plot.current = F,  -->
<!--                            plot.change = "relative", -->
<!--                            proj.names = proj.names) -->
<!--     ggsave(pl, file = paste0(out.dir, "6_mapchange-a2_lambda-rel.jpg"), height = 7, width = 9) -->

<!--     pl <- map_species_data(title = paste0(common, " (", sp.code, ")"),  -->
<!--                            region, -->
<!--                            plot = "psi",  -->
<!--                            out = out,  -->
<!--                            plot.current = F, -->
<!--                            plot.change = "relative", -->
<!--                            proj.names = proj.names) -->
<!--     ggsave(pl, file = paste0(out.dir, "6_mapchange-b2_psi-rel.jpg"), height = 7, width = 9) -->
<!--   } -->
<!-- } -->





<!-- # Validate ---- -->

<!-- # Get all data -->
<!-- species.data <- load_species_data(sp.code, -->
<!--                                   sp.code.all, -->
<!--                                   file.name = allfiles$file, -->
<!--                                   file.label = allfiles$name, -->
<!--                                   file.path = "DATA SWAMP/data-ready/", -->
<!--                                   keep.cols = covariates, -->
<!--                                   region = region,  -->
<!--                                   filter.region = filter.region, -->
<!--                                   year.start = year.start, -->
<!--                                   year.end = year.end, -->
<!--                                   coordunc = coordunc, -->
<!--                                   coordunc_na.rm = coordunc_na.rm, -->
<!--                                   spat.thin = spat.bal, -->
<!--                                   keep.conus.grid.id = gridkey$conus.grid.id) -->



<!-- all.auc <- get_AUC(species.data, out) -->



<!-- # Save everything ---- -->
<!-- if (block.out == "none") { -->
<!--   blockname <- "full" -->
<!--   save(region, file = paste0(out.dir, "region.rdata")) -->
<!-- } else {blockname <- block.out} -->

<!-- save(species.data, covar, covar_unscaled, data, constants, gridkey, all.auc, block.out, -->
<!--      file = paste0(out.dir, "data", blockname, "-info.rdata")) -->


<!-- # Final CV Figures ---- -->
<!-- # Make final cross validation figs if all blocks have been done -->

<!-- done <- list.files(path = out.dir) -->

<!-- mod.name <- c('data1','data2','data3','datafull') -->


<!-- if (all(c(paste0(mod.name, '.rdata'), paste0(mod.name, '-info.rdata')) %in% done)) { -->

<!--   # Load data from all model runs -->
<!--   auc <- c() -->
<!--   proc <- c() -->
<!--   obs <- c() -->
<!--   alpha <- c() -->
<!--   for (i in 1:3) { -->
<!--     load(paste0(out.dir, "data", i, "-info.rdata")) -->
<!--     load(paste0(out.dir, "data", i, ".rdata")) -->

<!--     # If validation data is missing, skip -->
<!--     if (length(all.auc) == 1) next -->

<!--     out$process.coef$block.out <- as.character(out$process.coef$block.out) -->
<!--     out$obs.coef$block.out <- as.character(out$obs.coef$block.out) -->
<!--     out$alpha$block.out <- as.character(out$alpha$block.out) -->

<!--     auc <- bind_rows(auc, all.auc) -->
<!--     proc <- bind_rows(proc, out$process.coef) -->
<!--     obs <- bind_rows(obs, out$obs.coef) -->
<!--     alpha <- bind_rows(alpha, out$alpha) -->

<!--   } -->

<!--   # If validation data exists... -->
<!--   if (!is.null(auc)) { -->

<!--     # now add full model -->
<!--     load(paste0(out.dir, "datafull.rdata")) -->

<!--     out$process.coef$block.out <- as.character(out$process.coef$block.out) -->
<!--     out$obs.coef$block.out <- as.character(out$obs.coef$block.out) -->
<!--     out$alpha$block.out <- as.character(out$alpha$block.out) -->

<!--     proc <- bind_rows(proc, out$process.coef) -->
<!--     obs <- bind_rows(obs, out$obs.coef) -->
<!--     alpha <- bind_rows(alpha, out$alpha) -->


<!--     load(paste0(out.dir, "datafull-info.rdata")) -->
<!--     auc$block <- as.character(auc$block) -->
<!--     auc <- bind_rows(auc, all.auc) -->


<!--     ggsave(plot_pars(out = proc,  -->
<!--                      plot.type = "cv",  -->
<!--                      plot.group = "process",  -->
<!--                      title = "Process parameter estimates", -->
<!--                      cov.labs = cov.labs), -->
<!--            file = paste0(out.dir, "7_CV-a-parameters-process.jpg"), height = 6, width = 10) -->

<!--     ggsave(plot_pars(out = alpha,  -->
<!--                      plot.type = "cv",  -->
<!--                      plot.group = "dataset",  -->
<!--                      title = "Dataset parameter estimates", -->
<!--                      cov.labs = cov.labs), -->
<!--            file = paste0(out.dir, "7_CV-b-parameters-dataset.jpg"), height = 6, width = 10) -->

<!--     ggsave(plot_pars(out = obs,  -->
<!--                      plot.type = "cv",  -->
<!--                      plot.group = "observation",  -->
<!--                      title = "Observation parameter estimates", -->
<!--                      cov.labs = cov.labs), -->
<!--            file = paste0(out.dir, "7_CV-a-parameters-observation.jpg"), height = 6, width = 10) -->


<!--     # Plot AUC -->
<!--     ggsave(plot_auc(auc, lines = T), file = paste0(out.dir, "7_CV-d_AUClines.jpg"), height = 7, width = 10) -->
<!--     ggsave(plot_auc(auc, lines = F), file = paste0(out.dir, "7_CV-d_AUCnolines.jpg"), height = 7, width = 10) -->
<!--   } -->
<!-- } -->


<!-- # Create time-summary output file ---- -->
<!-- cat("Summarization and visualization complete \n\n") -->

<!-- end3 <- Sys.time() - start3 -->

<!-- # Start output file -->
<!-- sink(paste0(out.dir,'time-summary_',block,'.txt')) -->

<!-- cat(paste0("Time summary for: ", nums.do,' ',sp.code,' ',model,'\n\n')) -->

<!-- cat(paste0('The model was run on ', format(Sys.time(), "%a %b %Y %d %X"),'\n\n')) -->


<!-- # Print total time to output file -->
<!-- cat('Total time for setup:\n') -->
<!-- print(end1) -->
<!-- cat('\n') -->

<!-- end2 <- readRDS(paste0(out.dir, 'time2_',block,'.rds')) -->
<!-- cat('Total time for fitting NIMBLE model: \n') -->
<!-- print(end2) -->
<!-- cat('\n') -->

<!-- cat('Total time for summarization: \n') -->
<!-- print(end3) -->
<!-- cat('\n') -->

<!-- cat('Start to finish time (code): \n') -->
<!-- print(end1 + end2 + end3) -->
<!-- cat('\n') -->

<!-- cat('Start to finish time (user delay): \n') -->
<!-- print(Sys.time() - start1) -->
<!-- cat('\n\n') -->

<!-- # Close output file -->
<!-- sink() -->

<!-- cat("Removing intermediate files...\n") -->
<!-- # system(paste0("rm ./outputs/03-species-models/MVPv1/",number,'_',sp.code,'_',model,"/setup_",block,"*")) -->
<!-- # system(paste0("rm ./outputs/03-species-models/MVPv1/",number,'_',sp.code,'_',model,"/samples_",block,"*")) -->
<!-- # system(paste0("rm ./outputs/03-species-models/MVPv1/",number,'_',sp.code,'_',model,"/time2_",block,".rds")) -->

<!-- cat(paste0("Model ",number,' ',sp.code,' ',model," complete!!\n")) -->


<!-- # End script -->


## High performance computing

We almost exclusively run these models on a high performance computing (HPC) cluster. We clone our repository directly to the HPC to ensure we have the same file structure and no version conflicts. We've set up the code so we can seamlessly move between running the workflow locally and on our HPC. 

To run the first steps of the workflow via `01-flexiSDM.R`, we call the script `01-flexiSDM.sh`, which takes 3 inputs: 

- Model number
- Block (defined by an array in the script)
- Local (1/0)

We have simplified this further by writing a wrapper called `01-HPC.sh`, which only takes one input: model number. Within the script, `local` is set to 0, indicating a run on an HPC.

The inputs provided to the `.sh` scripts are converted to inputs to the `01-flexiSDM.R` script and are interpreted below:

```{r hpc, eval = F}
# Converts command arguments to something interpretable by R
args = commandArgs(trailingOnly = TRUE)

if (length(args) > 0) {

  # Model number to run
  nums.do = as.numeric(args[1])

  # Blocks can be run as single jobs or as arrays
  block = as.numeric(args[2])
  
  # The block input is numeric only. Convert 4 = 'none'
  if (block == 4) {
    block <- 'none'
  }

  # Running locally? Yes = 1
  local = as.numeric(args[3])

  # If running on an HPC, set the working directory to the project home directory
  if (local == 0) {
    setwd('/home/directory/')
  }
}
```
